---
title: "code_assignment2"
date: "2025-05-29"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Assignment 2

```{r}
library(readr)   
library(effsize)     
library(pwr)  
library(lmtest)

```

```{r}
crime_data <- read_csv("crime.csv")
head(crime_data)

summary(crime_data)
```
## V)

```{r}
reg_crime <- lm(arrests ~ treated, data = crime_data)
summary(reg_crime)
```

```{r}
before_change <- crime_data$arrests[crime_data$treated == "before"]
after_change <- crime_data$arrests[crime_data$treated == "after"]
```

```{r}
t_test_result <- t.test(before_change, after_change, alternative = "greater")
t_test_result
```


## VI)Example: Metal detector at an airport

**Type I Error (False Positive)**:  
The alarm goes off, but the passenger has **no weapon**.  
 - We wrongly suspect a threat that doesn't exist.  
 - In hypothesis terms: we rejected the null (safe) when it was actually true.

**Type II Error (False Negative)**:  
The passenger **has a weapon**, but the detector does **not go off**.  
 - We fail to detect a real threat.  
 - We failed to reject the null (safe) when it was actually false.

## VII) Cohen's d effect size:

```{r}
d_size <- cohen.d(before_change, after_change, pooled=TRUE)
d_size
```

## VIII) Power Level
```{r}
table(crime_data$treated)

d_value <- as.numeric(d_size$estimate)
d_value
```

```{r}
pwr_result <- pwr.t.test(
  d = d_value, 
  n = 50, 
  sig.level = 0.05,
  type = "two.sample", 
  alternative = "two.sided")

pwr_result
```

We calculated the power of the two-sample t-test using the estimated Cohen’s d of 0.466 and a sample size of 50 per group.

Power is the probability that our test will correctly dtect a real effect or reject the null hypothesis when
the alternative is true.

The power result is around 64%. This means that if there really is an effect of this size, our test has a 64% chance of detecting it,
which is relatively low or moderate but not ideal, because it tends to lead to **Type 2 Errors (false negatives)**.


